
\section{Study Area}
% Provide a description of the sampling methods, field area, and geology.
% - Include maps or figures if necessary.

% want table with catchment characteristics
% also remember to include a table with abbreviations


The Melamchi-Indrawati catchment (85.441 - 85.601 E, 27.822 - 28.157 N) study area ranges from 790 to 5700 m a.s.l. (metres above sea level). The Melamchi River is a tributary of the larger Indrawati River and runs through the catchment draining an area of 325km$\ttmath{^2}$. 

\subsection{Geology and Geomorphic setting}

% Need to add Himalayan background

The geology of Melamchi is characterised by the characteristic banded gneiss, feldspathic schist and laminated quartzite of the Higher Himalayan Crystalline Sequence (HHCS). To the south of the confluence of the Melamchi River to its parent Indrawati river lies the Main Central Thrust (MCT) which separates the HHCS from the Lower Himalayan Sequence (LHS) (Dhital et al, Graf et al). The overall geology is therefore largely comprised of silicate metamorphic rock.

\bsk
\subsection{Climate and Vegetation}


There are two main climatic influences in the Himalayas: the monsoon system and the westerlies (Bookhagen and Burbank, 2010). The westerly winds typical of this latitude are responsible for the dry season in the Himalayas. The source of precipitation during the Indian Summer Monsoon (ISM) affecting Melamchi is the Bay of Bengal, due to the strong pressure gradient that changes the westerly winds to southerly winds. This is supported by isotopes... (Acharya et al, 2020). This temperature gradient reverses in the winter, when the oceans are warm and the High Himalaya is cold.  The Melamchi Khola catchment receives over 80\% of its rainfall during the monsoon. The seasonal variation in rainfall is thought to relate to different hydrological regimes, whereby river discharge and precipitation are 'coupled' when there is a significant enough amount of water to recharge the groundwater system. (Illien et al, 2021) The seasonal variation in precipitation therefore also translates to a variation in runoff, whereby this is twelve times stronger during the monsoon than during the dry season (Sharma, 1997). 

\bsk

\subsection{The effect of the monsoon on the landscape}
During the monsoon, higher elevations receive more water from the southerly winds compared to the westerlies. Oxygen isotopes suggest most of the precipitation occurs in the higher elevation parts of the catchment (Acharya et al, 2020), though lapse rate has been used to suggest the foothill slopes receive the most amount of rain (Kattel et al, 2012). The southern Himalayas are characterised by a large topographic gradient. This corresponds to a large temperature gradient contributing to tropical and alpine climates close to one another (Kattel et al, 2012). Changes in climate contribute to changes in the monsoonal system dynamics. The start of the monsoon has not changed in Nepal, but the end has been delayed. This has led to more intense precipitation on a per day basis, which is detrimental for crops in the winter season due to lack of moisture. Intense precipitation is also considered the main climatic cause of flooding (Panthi et al, 2015; Baniya et al, 2012). 


Annual mean temperatures in the Melamchi Khola Catchment range from ~24$^{\circ}$C at base elevation to ~8$^{\circ}$C at highest elevation sampled (3200 m a.s.l.). The area is characterised by a high erosion rate. "One-off" landslide events transport as much as four times the flux of sediment deposited in the valley in a year (Chan Mao Chen et al, 2023). These events are thought to be increasing in frequency over recent years as a result of climate change, increasing the erosion rate in these areas (Adhikari et al, 2023). In particular, effects of a flash flood in 2021 are still visible in the area, with damage done to several bridges and hundreds of families. (insert picture of flooding)

% Chan Mao Chen et al, 2023 calculate "Our sediment budget from DEM differencing indicates that the total erosion along the river was
% 83.24 million m3, whereas deposition in the Melamchi Valley was only 23.24 million m3". eg






% \begin{itemize}

% \item{What is the lat long extent of the study area?}

% 85.441 - 85.601 E, 27.822 - 28.157 N

% \item{What is the elevation range of the catchment?}

% 786 - 5697m

% \item{What is the area that the river drains? How do you calculate that (i.e. DEM)?}

% 325 km$^2$

% \item{What is the geology of the area?}



% \item{What is the climate of the area? what is it influenced by? i.e. monsoon, etc.}

% Bookhagen and Burbank (2010) identify two main climatic influences in the Himalayas: the monsoon system and the westerlies. The westerly winds, typical of this latitude, are responsible for the dry season in the Himalayas. The system is further divided into the East Asian and Indian Monsoon systems, which interact with each other. The high elevation of the High Himalayas creates a barrier that affects atmospheric circulation.  Bookhagen et al. (2005b) suggest that the Tibetan Plateau's high elevation generates a low-pressure cell near the surface, altering atmospheric circulation patterns.  This is one explanation for the monsoon, though other studies present differing views (see Bookhagen and Burbank, 2010 for a discussion). The source of precipitation during the Indian Summer Monsoon (ISM) affecting Melamchi is the Bay of Bengal, due to the strong pressure gradient that changes the westerly winds to southerly winds. This temperature gradient reverses in the winter, when the oceans are warm and the High Himalaya is cold.


% The high elevation of the High Himalayas creates a barrier that affects atmospheric circulation.  Bookhagen et al. (2005b) suggest that the Tibetan Plateau's high elevation generates a low-pressure cell near the surface, altering atmospheric circulation patterns.  This is one explanation for the monsoon, though other studies present differing views (see Bookhagen and Burbank, 2010 for a discussion). 


% Factor of 12 variation in runoff Tipper et al, 2006 and Sharma 1997


% \item{What is past literature on rainfall in monsoon season?}

% \item{What about clouds and fog? Especially when you are here?}

% \item{What are the annual mean temperatures at different elevations?}

% 24C at base elevation, ~900masl and 8C at highest elevation, ~3200masl.

% \item{What is the lapse rate like?}

% \item{What is the vegetation like?}

% Veget


% \item{Make sure to insert a description of the catchment: catchment, Area, mean slope, mean elevation, elevation range, land cover type, geology, Lat Long range}

\begin{table}[h!]
\centering
\begin{small}
\begin{tabular}{p{2cm} p{1cm} p{1cm} p{1.8cm} p{1.9cm} p{1.5cm} p{1.5cm} p{2.8cm}}
\hline
\textbf{Catchment} & \textbf{Area} & \textbf{Mean Slope} & \textbf{\phantom{hi}Mean Elevation} & \textbf{\phantom{}Elevation \phantom{hi}Range} & \textbf{\phantom{hi}NDVI \phantom{h}during} & \textbf{Geology} & \textbf{\phantom{hiidii}Location \phantom{oideo}Range} \\ 
                   & (km\(^2\))    & \phantom{h}(\%)                & \phantom{hdii}(m)                     & \phantom{hifi}(m)                      &           \textbf{Sampling}              &                   & \phantom{hiiiiiiiiii}(DD)             \\ \hline
       Melamchi            &     \phantom{h}325          &         \phantom{hi}20            &            \phantom{hdi}2400             &    786 - 5697                     &                         &     \phantom{hi}HHCS      &      85.441 - 85.601 E                  \\
       \phantom{hi}Khola         &               &                     &                         &                          &                         &                  &      27.822 - 28.157 N                 \\ \hline
\end{tabular}
\end{small}
\caption{Catchment characteristics of the study area.}
\label{tab:catchment_characteristics}
\end{table}


% elevation range from DEM
% area from DEM, and software to figure out catchment area

% \end{itemize}